### 2010-09-04 Spatial_adaptation/standing-variation.tex
/* Standing variation */  adjustments and constant-speed beginning

\documentclass{article}
\usepackage{amsmath,amssymb}

\newcommand{\var}{\mathop{\mbox{Var}}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\one}{\mathbf{1}}

\begin{document}

\section{Size of inflorescences}

Let the local rate of growth at location $x$ and time $t$ be given by $f(x,t)$,
and for each $x$ and $t$ let $(X^{(x,t)}_{s})_{s \ge t}$ have the distribution of a branching process begun with one individual at $x$ and time $t$,
and let $\kappa^{(x,t)}_s(dz) = \P\{ |X^{(x,t)}_s| \in dz \}$  be the marginal distribution of the size of $X^{(x,t)}$.
For a given mutation rate $\mu$ and population density $\rho(dx)$ we have a Poisson process of inflorescences of rate $\mu \rho$,
each of which having distribution $\kappa$.  
Then the set of inflorescences alive at time $t$ is Poisson, 
and the mean measure of such inflorescences is
\[ \nu(dx,A) = \mu \rho(dx) \int_{-\infty}^t \P\{ X^{(x,s)}_t \in A \} ds .\]
The first thing we'd want to do is to ignore the spatial extent of $X$ and only look at the total number of individuals,
so we need to know
\[ \int_{-\infty}^t \kappa^{(x,s)}_t(dz) ds . \]

Now assume everything is homogeneous, and let $Z_t$ be a continuous-time branching process begun at zero 
and having branching rate $1$ and offspring distribution $\{p_k : k\ge 0\}$.
Suppose that $Z$ has total lifetime $\tau$ and that $\tau$ is finite almost surely.
If we treat each inflorescence as a point mass, then the set of (location, mass) coordinates $(x,z)$
are Poisson with intensity $dx \otimes q(1,z) dz$ where $q(1,z)=q(z)$ is the Green function of $Z$, namely
\[  
  q(y,z) = \E^y \left[ \int_0^\tau \one(Z_t = z) dt \right] = \int_0^\infty P^t_{yz} dt .
\]
and $q$ solves $L q(y,z) = y \sum_k p_k (q(y+k-1,z)-q(y,z)) = \delta_{y=z}$.

If we are interested in the density of ultimately successful influorescences,
we need to futhermore weight by the probability that the offspring of any of the initial (assumed to be small) number of individuals in an inflorescence survive.
If $p_s$ is the probability of survival in the new selective regime,
this leads to the mean measure for successful influorescences
\[
   dx \int_0^\infty F(p_s,t) dt ,
\]
where $F$ is the generating function for $Z_t$,
\[
  F(a,t) = \E[a^{Z_t}] ,
\]
and let $\phi$ be the generating function for the offspring distribution,
\[
  \phi(a) = \sum_k p_k a^k .
\]
Now, $p_s$ is the minimal solution to $\phi(a) = a$,
and more generally, $F$ solves
\[
   \partial_t F(a,t) = \phi(F(a,t)) - F(a,t).
\]

\section{Properties of the mixed model}

Suppose that there are (point masses of) mutations present at time 0 as a Poisson process of rate $\lambda_0$ (calculated above),
and that after time 0 mutations arise at rate $\lambda$.
Suppose that the mutations spread distance $f(t)$ in time $t$, 
and that the volume of the cone of height $t$ is given by $h(t) = \int_0^t \omega_d f(s)^d ds$.

Let $\tau$ denote the time until the origin is adapted.
Then
\begin{align}
   \P \{ \tau > t \} &= \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\}, \quad \mbox{and} \\
   \E [ \tau ] &= \int_0^\infty \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\} .
\end{align}
As before, the mean density $\nu$ is equal to $1/\E[\tau]$.

Let $z_0$ be the mean proportion of space covered by mutations present at time 0.
Then, by the general result on competing inhomogeneous exponentials,
\begin{align}
    z_0 &= \P\{ \mbox{origin reached by mutations from $\lambda_0$ before $\lambda$} \} \\
        &= \int_0^\infty \lambda_0 d \omega_d f'(t) f(t)^{d-1} \exp \left\{ - \lambda_0 \omega_d f(t)^d - \lambda h(t) \right\} dt .
\end{align}
The mean density of new mutations that would have taken off but were prevented by standing variation is
\begin{align}
    \nu_* = \int_0^\infty \lambda \exp\{ - \lambda h(t) \} \left( 1 - \exp\{ -\lambda_0 \omega_d f(t)^d \} \right) dt
\end{align}
and so the mean proportion of new mutations that would be successful in the absence of standing variation, but are prevented by standing variation,
is 
\begin{align}
    z_* = \frac{ \int_0^\infty \lambda \exp\{ - \lambda h(t) \} \left( 1 - \exp\{ -\lambda_0 \omega_d f(t)^d \} \right) dt }
                { \int_0^\infty \lambda \exp\{ - \lambda h(t) \} dt } .
\end{align}

If we rescale time by $\alpha$ and space by $\beta$, 
then this transforms $(\lambda_0, \lambda, f(\cdot))$ into $(\lambda_0/\beta^d, \lambda/(\alpha \beta^d), f(\cdot/\alpha)/\beta)$.  
Setting $\beta = \lambda_0^{1/d}$ and $\alpha = \lambda / \lambda_0$, this is mapped to $(1,1,f(\cdot \lambda_0/\lambda)/\lambda_0^{1/d})$.

\subsection{Constant speed}

Now let $f(t) = vt$, so $h(t) = \frac{ \omega_d v^d }{ d+1 } t^{d+1}$.
The integrals that appear above are of the following form:
\begin{align}
  \int_0^\infty t^c \exp \left( - \alpha t^a - \beta t^b \right) dt 
            &= \left( b^{-1} \beta^{ (1-c)/b } \right) \int_0^\infty u^{(c+1-b)/b} \exp\left( - \alpha \beta^{-a/b} u^{a/b} - u \right) du ,
\end{align}
so it suffices to learn something about the function
\begin{equation}
    G(a,b,x) := \int_0^\infty  t^a \exp\left( -x t^b - t \right) dt .
\end{equation}
We will only be interested in this as a function of $x$; 
the coefficients $a$ and $b$ will merely depend on the dimension and what quantity we are calculating.
We can construct a power series for $G$ in $x$:
\begin{align}
    \partial_x^n G(a,b,x) &= \int_0^\infty (-1)^n t^{a+nb} \exp\left( -x t^b - t \right) dt \\
            &= (-1)^n G(a+nb,b,x) , \quad \mbox{so} \\
    \partial_x^n G(a,b,x) \vert_{x=0} &= \Gamma(a+nb+1) .
\end{align}
So, if everything works out, we will have that
\begin{align}
    G(a,b,x) = \sum_{n \ge 0} \frac{(-x)^n}{n!} \Gamma(a+nb+1) .
\end{align}

A few facts, first.
Using repeatedly the identity $\Gamma(x+1) = x\Gamma(x)$
and the Pochammer symbol $(x)_n = x(x+1)\cdots(x+n-1)$,
note that for $m \in \N$,
\[
    \Gamma(m+y) = (y)_m \Gamma(y).
\]
Also note that for $c$ and $m$ integers,
\[
    (y)_{cm} = c^{cm} \prod_{k=0}^{c-1} (\frac{y+k}{c})_{m} ,
\]
and in particular, that
\[
    (cm+\ell)! = \ell! (\ell+1)_{cm} = \ell! c^{cm} \prod_{k=0}^{c-1} (\frac{\ell+1+k}{c})_m .
\]

Now suppose that $b$ is rational (as will be the case), letting $b = b'/c$ with $b'$ and $c$ positive integers.
Therefore, setting $n = mc+\ell$,
\begin{align}
    G(a,b,x) &= \sum_{\ell=0}^{c-1} \sum_{m \ge 0} \frac{(-x)^{mc+\ell}}{(mc+\ell)!} \Gamma(a+\ell b'/c+mb'+1) \\
            &= \sum_{\ell=0}^{c-1} \sum_{m \ge 0} \frac{(-x)^{mc+\ell}}{(mc+\ell)!} (a+\ell b')_{mb'} \Gamma(a + \ell b') \\
            &= \sum_{\ell=0}^{c-1} \frac{ (-x)^{\ell} }{ \ell! } \Gamma(a + \ell b') \sum_{m \ge 0} ( (-cx)^c / (b')^{b'} )^m 
                    \frac{ \prod_{j=0}^{b'-1} (\frac{ a+b'\ell+j }{ b' } )_m }{ \prod_{k=0}^{c-1} (\frac{\ell+1+k}{c})_m } ,
\end{align}
which is the sum of $c$ (generalized) hypergeometric series.
For these to converge, we need $b' \le c$, e.g.\ $b\le 1$.

In particular, if we let $A=\lambda_0 \omega_d v^d $ and $B=\lambda h(1)$, then 
\begin{align}
    \E[\tau] &= \int_0^\infty e^{-A t^d - B t^{d+1} dt \\
            &= \frac{B^{1/(d+1)}}{d+1} \int_0^\infty u^{-d/(d+1))} \exp\left( -AB^{-d/(d+1))} u^{d/(d+1)} -u \right) du \\
            &= \frac{B^{1/(d+1)}}{d+1} G \left(-d/(d+1), d/(d+1), A B^{-d/(d+1)} \right) .
\end{align}

\section{Poisson--Vorono\u\i\ facts}

Fill in some facts about PV tesselations.

\end{document}