### 2011-04-14 Spatial_adaptation/local-adaptation.tex
/* Local Adaptation */ more notes

\documentclass{article}
\usepackage{amsmath,amssymb}

\newcommand{\var}{\mathop{\mbox{Var}}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\one}{\mathbf{1}}

\begin{document}

Suppose that mutants have selective (dis-)advantage $s(x)$ at location $x$,
that the symmetric dispersal kernel is $\kappa(\cdot)$, and that the mean-squared dispersal distance is $\sum_y y^2 \kappa(y) = \sigma^2$.
Then in the large population limit, if $\phi(t,x)$ is the proportion mutants at location $x$ at time $t$,
if $s$ and $\sigma^2$ are of the same (small) order,
(see Peter's slides from Berkeley probability seminar)
\begin{align}
    \partial_t \phi(t,x) &\approx \sum_y s(x) \kappa(y-x) (1-\one_{x=y}) \phi(t,y)(1-\phi(t,x)) 
            + \sum_y \kappa(y-x) (1-\frac{1}{N}\one_{x=y}) (\phi(t,x)-\phi(t,y))  \\
        &\approx s(x) \phi(t,x) (1-\phi(t,x)) + \frac{1}{2} \sigma^2 \partial_x^2 \phi(t,x) .
\end{align}

Note that if $\tau$ is the first time that the new patch $A$ is occupied,
and $Z_t$ is the state of the proces, then
\[
    \partial_t \P\{ \tau > t \} = \int \int_A s(x) \kappa(y-x) \E[ Z_t(dx) | \tau > t ] dy \P\{ \tau > t \}
\]
... so we want to find the first integral (and show that for $t>\epsilon$ it is approximately constant).
Since we are also taking the small-dispersal limit, this seems tricky.

Another option: we are trying to find the first time that a mutant appears in patch $A$ {\em and fixes locally};
if the benefit is small then there may be many unsuccessful migrants first.
Suppose we modify the process so that offspring appearing in $A$ have distribution {\em conditioned on local extinction};
then we suppose that each such individual has probability ${}\approx 2s_b$ of fixation,
so that the mean number of such offspring per generation times $2s_b$ gives us the rate we are looking for.
A supercritical branching process conditioned on extinction is a subcritical branching process,
with mean offspring number $\partial_z \E[z^N]\vert_{z=p}$, where $N$ is the offspring number and $p$ is the probability of extinction.
In the case of a binomial (check this?) branching, this changes $s$ to $-s$.

\end{document}