### 2012-12-03 maize-ex.tex
WorkingWiki replaced project description

\documentclass{article} \usepackage[landscape]{geometry} \begin{document}

\input{helianthus-ex-table.tex}

\end{document}