### 2013-01-31 branching-transit.tex
WorkingWiki replaced project description

\documentclass{article}
\usepackage{fullpage} 
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\var}{\mathop{\mbox{Var}}}

\begin{document}

Consider a largeish occupied patch of habitat in which an allele is beneficial ($s_b$),
surrounded by habitat where it is deleterious ($s_d$).
The allele is present in surrounding areas at low frequency due to migration-selection balance.
We want the first time that another patch at distance $x$ is successfully colonized.

The general picture:
\begin{itemize}

\item The excursions sufficiently far from the patch are like subcritical branching processes,
with mean $1-s_d$ and variance $\xi^2$.

\item An excursion that makes it to the new patch in $k$ copies establishes with probability $1-(1-p_s)^k$,
where $p_s \approx 2s_b/\xi^2$, so establishment has approximate probability $2 k s_b / \xi^2$.

\item The probability that the branching process lives for $t$ generations is asymptotically a constant times $(1-s_d)^t$.

\item Conditioned on living for time $t$, the process can be represented by a single lineage whose offspring distribution
is the original distribution, conditioned to be nonzero, minus one.
(Also, it is known that conditional on being alive at $t\to\infty$, the allele is present in $K$ copies; where $K$ has calculable mean and variance.)

\item Spatial motion is independent of branching;
so we can take th emotion of the trunk lineage to be Brownian motion;
the chance that the motion first hits $x$ at time $t$ is (Borodin \& Salminen 2.02)
\[
\frac{x}{t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t}\right) .
\]

\item Therefore, the probability that an excursion makes it to $x$ at time $t$ is
\[
  (1-s_d)^t \frac{x}{t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t}\right) ,
\]
or the chance that it is still alive then time the chance the trunk lineage hits $x$ at that time.

\item This induces a distribution on $T$, the time that a successful excursion took to cross,
and hence the length of the hitchhiking segment.

\end{itemize}

From Jagers, section 2.6,
the mean and variance of $K$ are as follows.
Let the probability of $k$ offpsring be $p_k$ and let
\[
 f(u) =  \sum_k p_k u^k,
\]
and $f_n$ the $n^\mathrm{th}$ composition of $f$.
Note that the mean number of offspring is $1-s_b = f'(1)$, and the variance is $\xi^2 = f''(1) + f'(1)(1-f'(1)) = f''(1) + s_d(1-s_d)$.
If $\sum_k p_k k \log k < \infty$, then 
\[
\phi = \lim_{n \to \infty} \frac{ 1-f_n(0) }{ (1-s_d)^n } > 0 .
\]
Furthermore,
\begin{gather*}
\E[K] = 1/\phi \qquad \mbox{and} \\
\E[K(K-1)] = \frac{ f''(1) }{ \phi s_d (1-s_d) } \qquad \mbox{so} \\
\var[K] = \frac{1}{\phi} \left( \frac{ \xi^2 }{ s_d (1-s_d) } + \frac{1}{1-\phi} - 1 \right) .
\end{gather*}

\end{document}